---
title: "data_gathering"
author: "Patrick Gyasi"
date: "2023-09-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data gathering for a Bayesian linear regression model of the NKPC
MA Thesis, Summer 2023                                  
Objective: estimate the slope of the New Keynesian Phillips Curve for the 
EA-11 country composition in a pooled regression             
pi_t = a * pi_t+1 + b * x_t + e_t 


```{r}

library(tidyverse)

```


## Download data

### ECB Data


```{r}

country_names_ecb <- c("AT", "BE", "DE", "ES", "FI", "FR", "IE", "IT", "LU", "NL", "PT")

load_ecb_data <- function(str1, str2){
  
  # install.packages(c("ecb", "tidyverse"))
  
  load_ecb_data <- vector(mode = "list", length = length(country_names_ecb)) # load_ecb_data <- list()
  names(load_ecb_data) <- country_names_ecb
  
  for(i in names(load_ecb_data)) {
    load_ecb_data[[i]] <- ecb::get_data(
      paste0(str1, i, str2) 
    )
  }
  
  load_ecb_data <- do.call(rbind, load_ecb_data) |> # adds row names to the observations
    dplyr::select(ref_area, obstime, obsvalue) |> 
    dplyr::rename(
      country = ref_area,
      period  = obstime,
    ) |> 
    dplyr::arrange(period, country)
  
  return(load_ecb_data)
}

```


#### Inflation

##### HICP - All-items excluding energy and food, monthly annual rate of change
Eurostat, Neither seasonally nor working day adjusted  
[https://sdw.ecb.europa.eu/quickview.do?SERIES_KEY=122.ICP.M.AT.N.XEF000.4.ANR](https://sdw.ecb.europa.eu/quickview.do?SERIES_KEY=122.ICP.M.AT.N.XEF000.4.ANR)

```{r}
# ICP.M.AT.N.XEF000.4.ANR
hicpX_mon <- load_ecb_data("ICP.M.", ".N.XEF000.4.ANR") |> 
  filter(between(period, "1998-01", "2022-12")) |> 
  rename(hicpX_mon = obsvalue)

```


##### Unemployment rate
Total, Age 15 to 74; **seasonally adjusted**, not working day adjusted  
Unit:	Percent, quarterly  
[https://sdw.ecb.europa.eu/quickview.do?SERIES_KEY=477.IESS.Q.FI.S.UNEHRT.TOTAL0.15_74.T](https://sdw.ecb.europa.eu/quickview.do?SERIES_KEY=477.IESS.Q.FI.S.UNEHRT.TOTAL0.15_74.T)

```{r}
# IESS.Q.FI.S.UNEHRT.TOTAL0.15_74.T
unemp_rate_adj <- load_ecb_data("IESS.Q.", ".S.UNEHRT.TOTAL0.15_74.T") |> 
  filter(between(period, "1998-Q1", "2022-Q4")) |> 
  rename(unemp_rate_adj = obsvalue)

```


*GDP at current prices, calendar and seasonally adjusted not available*

Chained volume series (Chain Linking):
In a chain linked GDP series, we calculate the value of goods and services produced this year using the prices these things cost last year. In that way, we remove the effect of price (value effect) and the remaining GDP growth will be due to more goods and services being produced (volume effect). It is usually the volume effect that is the most meaningful. [Ireland's national statistical office CSO](https://www.cso.ie/en/interactivezone/statisticsexplained/nationalaccountsexplained/chainlinking/)

GDP at chained volume measure is a series of GDP statistics adjusted for the effect of inflation to give a measure of ‘real GDP’.[Economicshelp.org](https://www.economicshelp.org/blog/7397/economics/gdp-at-chained-volume-measure/)

##### Gross domestic product at market prices, quarterly
Domestic (home or reference area), Total economy   
Euro, **Chain linked volume** (rebased)  
Non transformed data, **Calendar and seasonally adjusted** data  
[https://sdw.ecb.europa.eu/quickview.do?SERIES_KEY=320.MNA.Q.Y.LU.W2.S1.S1.B.B1GQ._Z._Z._Z.EUR.LR.N](https://sdw.ecb.europa.eu/quickview.do?SERIES_KEY=320.MNA.Q.Y.LU.W2.S1.S1.B.B1GQ._Z._Z._Z.EUR.LR.N)

```{r}
# MNA.Q.Y.LU.W2.S1.S1.B.B1GQ._Z._Z._Z.EUR.LR.N
gdp_chain_adj <- load_ecb_data("MNA.Q.Y.", ".W2.S1.S1.B.B1GQ._Z._Z._Z.EUR.LR.N") |> 
  filter(between(period, "1998-Q1", "2022-Q4")) |> 
  rename(gdp_chain_adj = obsvalue)

```


##### Gross domestic product at market prices, quarterly
Domestic (home or reference area), Total economy  
Domestic currency (incl. conversion to current currency made using a fixed parity)  
*Chain linked volume* (rebased), **Growth rate, over 1 year**  
*Calendar and seasonally adjusted* data  
[https://sdw.ecb.europa.eu/quickview.do?SERIES_KEY=320.MNA.Q.Y.IE.W2.S1.S1.B.B1GQ._Z._Z._Z.XDC.LR.GY](https://sdw.ecb.europa.eu/quickview.do?SERIES_KEY=320.MNA.Q.Y.IE.W2.S1.S1.B.B1GQ._Z._Z._Z.XDC.LR.GY)

```{r}
# MNA.Q.Y.IE.W2.S1.S1.B.B1GQ._Z._Z._Z.XDC.LR.GY
gdp_growth_yoy_adj <- load_ecb_data("MNA.Q.Y.", ".W2.S1.S1.B.B1GQ._Z._Z._Z.XDC.LR.GY") |> 
  filter(between(period, "1998-Q1", "2022-Q4")) |> 
  rename(gdp_growth_yoy_adj = obsvalue)

```


### OECD Data

#### expected inflation

[OECD Inflation forecast](https://data.oecd.org/price/inflation-forecast.htm)
[Inflation forecast quarterly](https://data.oecd.org/chart/77YT)

```{r}

link_exp_inf <- "https://stats.oecd.org/sdmx-json/data/DP_LIVE/AUT+BEL+DEU+ESP+FIN+FRA+IRL+ITA+LUX+NLD+PRT.CPIFORECAST.TOT.AGRWTH.Q/OECD?contentType=csv&detail=code&separator=comma&csv-lang=en&startPeriod=1998-Q1&endPeriod=2022-Q4"

exp_infl <- read_csv(link_exp_inf) |> 
  select(LOCATION, TIME, Value) |> 
  rename(
    country  = LOCATION,
    period   = TIME,
    exp_infl = Value
  ) |> 
  arrange(period, country)

```


## Tidy data

standardizing country names

```{r}

country_names_oecd <- unique(exp_infl$country)  # replace with ecb country names
 
for(i in seq_along(country_names_oecd)) {
  exp_infl$country <- gsub(country_names_oecd[i], 
                           country_names_ecb[i], 
                           exp_infl$country
                           )
}

```

## compute quarterly data from monthly data for hicpX

```{r}

hicpX_list <- vector(mode = "list", length = length(country_names_ecb))
names(hicpX_list) <- country_names_ecb

for(i in names(hicpX_list)) {
  
  hicpX_geo_quart <- hicpX_mon |> 
    filter(country == i)
  
  hicpX_geo_quart$hicpX <- round(
    zoo::rollapply(hicpX_geo_quart$hicpX_mon, 
                   width = 3, FUN = mean,
                   align = "right", fill = NA),
    digits = 2)
  
  hicpX_geo_quart$period <- as.Date(paste0(hicpX_geo_quart$period, "-01"))
  hicpX_geo_quart$month  <- format(hicpX_geo_quart$period, format = "%m")
  
  hicpX_geo_quart <- na.omit(hicpX_geo_quart) |>
    filter(month %in% c("03", "06", "09", "12")) |>
    select(country, period, hicpX)  |> 
    arrange(period)
  
  hicpX_list[[i]] <- hicpX_geo_quart
  
}

hicpX <- do.call(rbind, hicpX_list) |> 
  arrange(period, country)

```

### merging data

```{r}
data_list <- list(hicpX, exp_infl, unemp_rate_adj, unemp_gap_adj,
                  gdp_chain_adj, gdp_growth_yoy_adj
                  ) |> 
  setNames(c("hicpX", "exp_infl", "unemp_rate_adj",
             "unemp_gap_adj", "log(gdp_chain_adj)", 
             "gdp_growth_yoy_adj")
           )

# standardize period format to YYYY-MM-DD
quart_char <- paste0("Q", 1:4)
quart_num <- c(paste0(0, c(3, 6, 9)), 12)

for(i in names(data_list)) {
  for(j in 1:4) { # standardize period format to YYYY-MM-DD
    data_list[[i]]$period <- gsub(quart_char[j], quart_num[j],
                                             data_list[[i]]$period
    )
  } # change periode type to date
  data_list[[i]]$period <- as.Date(
    paste0(data_list[[i]]$period, "-01")
  )
}

# join all quarterly data (keep NA's --> full_join)
data_NKPC <- data_list |> 
  reduce(full_join, by = c("period", "country"))

data_NKPC$log_gdp_chain <- log(gdp_chain_adj[, 3, drop = TRUE])
data_NKPC$gdp_chain_adj <- NULL

data_NKPC <- data_NKPC |> 
  relocate(log_gdp_chain, .before = gdp_growth_yoy_adj)
```

#### non-lagged variables
```{r}

data_NKPC_contemp <- data_list |> 
  reduce(full_join, by = c("period", "country"))

data_NKPC_contemp$log_gdp_chain <- log(gdp_chain_adj[, 3, drop = TRUE])
data_NKPC_contemp$gdp_chain_adj <- NULL

data_NKPC_contemp <- data_NKPC_contemp |> 
  relocate(log_gdp_chain, .before = gdp_growth_yoy_adj)

```


#### lag variables
```{r}

data_NKPC <- data_NKPC |>
  group_by(country) |> 
  mutate(
    unemp_rate_adj     = lag(unemp_rate_adj, n = 4),
    unemp_gap_adj      = lag(unemp_gap_adj, n = 4),
    log_gdp_chain      = lag(log_gdp_chain, n = 4),
    gdp_growth_yoy_adj = lag(gdp_growth_yoy_adj, n = 4)
  ) |> 
  ungroup() # |>  
# filter(period != "1998-09-01":"1998-12-01")

data_NKPC <- data_NKPC[45:1100, ]

```
















































